\Title

### 2021-06-04 latex/paper/main.tex
Finished writing proof of theorem, need to edit now.

\subsection{Energy gain on \(\mathcal{O}_1\)}
We will now prove the first part of Theorem
\ref{thm:acrobot-energy-stabilization}.
We choose \(D\) in Lemma \ref{lemma:energy-regulation-D} to be the
annulus \(\mathcal{O}_1 \backslash \{(0,0)\}\) consisting of all oscillations of
the nominal pendulum.
Using the procedure described by Mohammadi \etal 
\cite{dynamic_vhcs_stabilize_closed_orbits} for finding pseudo-polar
coordinates adapted to level sets of \(E\), we compute
\begin{align}
    \label{eqn:r-osc}
    r &= \arccos\left(\cos(q_u) - \frac{p_u^2}{30m^2gl^3}\right)
    , \\
    \label{eqn:theta-osc}
    \theta &=
    \left.\arctan_2\left(-\sign{p_u}\sqrt{1-\frac{q_u^2}{r^2}},\frac{q_u}{r}\right)
    \right|_{r = r(q_u,p_u)}
    ,
\end{align}
where \(r \in \, ]0,\pi[\).
We now verify these coordinates satisfy the requirements of 
Lemma \ref{lemma:energy-regulation-D}.

When \(I = 0\) we have \(\dot{r} = 0\) and
\begin{equation} \label{eqn:theta-dot-nom-osc}
    \dot{\theta} = \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(r\cos(\theta)) - \cos(r)}
            {r^2 \sin(\theta)^2}}
    .
\end{equation}
Note that ~\eqref{eqn:theta-dot-nom-osc} has removable singularities at
\(\theta \in \{0,\pi\}\).
Taking the limit as \(\theta\) approaches these points yields
\begin{equation}\label{eqn:theta-dot-limit}
    \lim\limits_{\theta \to 0}\dot{\theta} 
    = \lim\limits_{\theta \to \pi} \dot{\theta}
    = \sqrt{\frac{6g \sin(r)}{10lr}}
    , 
\end{equation}
which is smooth and well-defined for all
\(r \in \, ]0,\pi[\).
From \eqref{eqn:theta-dot-nom-osc}-\eqref{eqn:theta-dot-limit},
one can verify that \(\dot{\theta} > 0\) for every \(r \in \, ]0,\pi[\).
Defining \(g(\theta,r,I) = \dot{r}/\dot{\theta}\), MATLAB symbolic computations
confirm that \((\partial g/\partial r)(\theta,r,0) \equiv 0\) as required.

@@

\textbf{TODO: show orbits will go through Epi, 
    then derive the constrained dynamics for the rotation domain and apply lemma
    there, concluding that O2 is O1 U Epi U rotation domain and is therefore
    also gaining energy.}

>>>>

\subsection{Energy gain on \(\mathcal{O}_1\)}
We will now prove the first part of Theorem
\ref{thm:acrobot-energy-stabilization}.
We choose \(D\) in Lemma \ref{lemma:energy-regulation-D} to be the
annulus \(\mathcal{O}_1 \backslash \{(0,0)\}\) consisting of all oscillations of
the nominal pendulum.
Using the procedure described by Mohammadi \etal 
\cite{dynamic_vhcs_stabilize_closed_orbits} for finding pseudo-polar
coordinates adapted to level sets of \(E\), we compute
\begin{align}
    \label{eqn:r-osc}
    r &= \arccos\left(\cos(q_u) - \frac{p_u^2}{30m^2gl^3}\right)
    , \\
    \label{eqn:theta-osc}
    \theta &=
    \left.\arctan_2\left(-\sign{p_u}\sqrt{1-\frac{q_u^2}{r^2}},\frac{q_u}{r}\right)
    \right|_{r = r(q_u,p_u)}
    ,
\end{align}
where \(r \in \, ]0,\pi[\).
This has inverse
\begin{align*}
    q_u &= r\cos(\theta)
    , \\
    p_u &= -\sign{\sin(\theta)}
    \sqrt{30m^2gl^3(\cos(rc_\theta - c_r)}
    .
\end{align*}
We now verify these coordinates satisfy the requirements of 
Lemma \ref{lemma:energy-regulation-D}.

When \(I = 0\) we have \(\dot{r} = 0\) and
\begin{equation} \label{eqn:theta-dot-nom-osc}
    \dot{\theta} = \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(r\cos(\theta)) - \cos(r)}
            {r^2 \sin(\theta)^2}}
    .
\end{equation}
Note that ~\eqref{eqn:theta-dot-nom-osc} has removable singularities at
\(\theta \in \{0,\pi\}\).
Taking the limit as \(\theta\) approaches these points yields
\begin{equation}\label{eqn:theta-dot-limit}
    \lim\limits_{\theta \to 0}\dot{\theta} 
    = \lim\limits_{\theta \to \pi} \dot{\theta}
    = \sqrt{\frac{6g \sin(r)}{10lr}}
    , 
\end{equation}
which is smooth and well-defined for all
\(r \in \, ]0,\pi[\).
From \eqref{eqn:theta-dot-nom-osc}-\eqref{eqn:theta-dot-limit},
one can verify that \(\dot{\theta} > 0\) when \(I = 0\), for every 
\(r \in \, ]0,\pi[\).
Defining \(g(\theta,r,I) = \dot{r}/\dot{\theta}\), MATLAB symbolic computations
confirm that \((\partial g/\partial r)(\theta,r,0) \equiv 0\) as required.

@@

Before moving on to the second part of Theorem
\ref{thm:acrobot-energy-stabilization}, we must first confirm that any acrobot
constrained by \eqref{eqn:acrobot-constraint} will eventually start rotating.
The definition of energy gain states that all orbits initialized in
\(\mathcal{O}_1\) will escape compact subsets of \(\mathcal{O}_1\) in finite time. 
This is not enough to prove that all orbits will escape the boundary of
\(\mathcal{O}_1\).

Let \(E_\pi\) be the level set with energy \(E(\pi,0)\), 
which forms the boundary of \(\mathcal{O}_1\). 
Define \(\bar{\mathcal{O}}_1 = \mathcal{O}_1 \cup E_\pi\) to be the closure of
\(\mathcal{O}_1\).
An orbit of the acrobot will only begin rotating if it escapes
\(\bar{\mathcal{O}}_1\) by crossing through \(E_\pi\), and eventually remains
outside this set.
It is possible for orbits starting in \(\mathcal{O}_1\) to always approach
\(E_\pi\) without ever crossing into the rotation zone.
We will prove this does not happen in general.

To begin, let us analyze the upright equilibrium.
Taking the Jacobian of the constrained dynamics
~\eqref{eqn:acrobot-constrained-dynamics} at \((q_u,p_u) = (\pi,0)\) yields
\[
    J = \begin{bmatrix}
        -\frac{6mgl\bar{q}_aI}{5} & \frac{1 - 2m^2gl^3\bar{q}_a^2 I^2}{5ml^2} \\
        3mgl & mgl\bar{q}_aI
    \end{bmatrix}
    ,
\]
which has characteristic polynomial
\(\det\left(\lambda \Id{2} - J\right)
= \lambda^2 + 1/5 mgl\bar{q}_a I \lambda - 3g\).
By Descartes' rule of signs, this polynomial has one root with positive real
part. 
The equilibrium \((\pi,0)\) is therefore unstable, so the stable manifold \(\Pi^+\) of
initial conditions converging to \((\pi,0)\) is one-dimensional, and hence is
of measure zero in \(\SxR\).

Using \(x(t) := (q_u(t),p_u(t))\) as shorthand, let
\(x(0) \in \mathcal{O}_1\) be a nonzero initial condition of the acrobot.
Suppose by way of contradiction that the orbit \(x(\R)\) is confined within 
\(\bar{\mathcal{O}}_1\), and does not exit through \(E_\pi\) into the rotation
zone.
Since \(\bar{\mathcal{O}}_1\) is compact, the Birkhoff Theorem \cite{birkhoff}
implies:
\begin{itemize}
    \item The positive limit set \(L_+\) of \(x(t)\) is non-empty, compact, and
        invariant.
    \item The solution \(x(t)\) asymptotically tends to \(L_+\).
\end{itemize}
Since the acrobot gains energy on \(\mathcal{O}_1\), the positive limit set of
\(x(t)\) must be the largest invariant subset of \(E_\pi\). 
From our discussion on pseudo-polar coordinates (in particular the extension
\ref{eqn:theta-dot-limit} of \(\dot{\theta}\) to \(\bar{\mathcal{O}}_1\)),
we know that \(\dot{\theta} \geq 0\) on \(E_\pi\) with equality if and only if 
\(\theta \in \{0,\pi\}\).
Taking the derivative of \(E(q_u,p_u)\) at the \(p_u\)-axis yields
\(\dot{E}(0,p_u) = g/(5l) \sin(q_a)p_u\).
This is non-zero everywhere except at the origin, which means it is non-zero on
\(E_\pi\); this implies \(E_\pi\) is not invariant, so the positive limit set of 
\(x(t)\) must be the set \(\{(\pi,0)\}\). Hence, \(x(t)\) converges to 
\((\pi,0)\), so \(x(0) \in \Pi^+\).
Since \(\Pi^+\) is a set of measure zero in \(\SxR\), almost every orbit
initialized in \(\mathcal{O}_1\) must escape through \(E_\pi\) into the rotation
domain.

Escaping once into the rotation domain does not guarantee the acrobot
orbit will never return to \(\mathcal{O}_1\).
Indeed, the orbit could escape through \(E_\pi\) and return to \(\mathcal{O}_1\)
several times.
However, almost every orbit must eventually remain outside the closure of 
\(\mathcal{O}_1\). 
If not, the positive limit set would once again be
\(E_\pi\); by the same argument as above, that orbit would be in \(\Pi^+\).

We conclude that all orbits beginning in \(\mathcal{O}_1\) will, in finite time,
escape the closure of \(\mathcal{O}_1\) and remain in the rotation domain
forever after.

\subsection{Energy Gain on \(\mathcal{O}_2(\bar{\rho})\)}
Since \(\mathcal{O}_2(\bar{\rho}) = \bar{\mathcal{O}}_1 \cup \mathcal{R}\)
(where \(\mathcal{R}\) is the rotation domain defined in
\eqref{eqn:rotation-domain}), and since all orbits escape 
\(\bar{\mathcal{O}_1}\) in finite time, we need simply prove energy gain on
\(\mathcal{R}\) to complete the proof of Theorem
\ref{thm:acrobot-energy-stabilization}.
We separate \(\mathcal{R}\) into its two distinct connected components:
\(\mathcal{R}^+\), the set of rotations with \(p_u > 0\); 
and \(\mathcal{R}^-\), the set of rotations with \(p_u < 0\).
We take the domain of interest for Lemma \ref{lemma:energy-regulation-D} to be
\(D \in \{\mathcal{R}^+, \mathcal{R}^-\}\).

We now apply the method described in
\cite{dynamic_vhcs_stabilize_closed_orbits} to find pseudo-polar coordinates
adapted to rotations.
These coordinates are 
\begin{align}
    \label{eqn:r-rot}
    r &= \sqrt{p_u^2 - 30m^2gl^3(1-c_u)}
    , \\
    \label{eqn:theta-rot}
    \theta &= \pm q_u
\end{align}
where \(r \in \, ]\sqrt{60m^2gl^3}, \bar{\rho}[\) and \(\pm\) denotes
whether \(D=\mathcal{R}^+\) or \(D=\mathcal{R}^-\).
The inverse of this transformation is
\begin{align*}
    q_u &= \pm \theta
    , \\
    p_u = \pm \sqrt{r^2+30m^2gl^3(1-c_\theta)}
    .
\end{align*}

When we compute the constrained dynamics in
\((r,\theta)\) coordinates and set \(I = 0\), MATLAB symbolic computations
reveal that \(\dot{r} = 0\) and
\(\dot{\theta} = 1/(5ml^2) \sqrt{r^2 - 30m^2gl^3(1-c_\theta)}\).
Regardless of \(D\), we have \(\dot{\theta} > 0\) when \(I = 0\) as required.
Defining \(g(\theta,r,I) = \dot{r}/\dot{\theta}\), MATLAB symbolic computations
confirm that \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
The function \(b(\theta,r_0) = (\partial g/\partial I)(\theta,r_0,0)\) is given
in the statement of Theorem \ref{thm:acrobot-energy-stabilization}.
The theorem also assumes that \(S(r_0) \geq \epsilon\).
Thus, all requirements of Lemma \ref{lemma:energy-regulation-D} are satisfied,
so there exists some \(I_2^\star(\bar{\rho})\) whereby the acrobot gains/loses
energy on \(\mathcal{R}^+\) and \(\mathcal{R}^-\).

Let \(I_1^\star\) be the value returned by Lemma \ref{lemma:energy-regulation-D}
for \(\mathcal{O}_1\).
We conclude that the acrobot gains/loses energy on \(\mathcal{O}_2(\bar{\rho})\)
for \(|I| \leq I^{\star\star}\), where 
\(I^{\star\star} = \min \{I_1^\star, I_2^\star(\bar{\rho})\}\).
This completes the proof of Theorem \ref{thm:acrobot-energy-stabilization}.
